---
title: "Assignment 7"
format: 
    html:
        output-file: "index.html" 
---

## Problem 1:

```{r}
#|  echo: false
#|  warning: false
library(dplyr)
library(tidyr)
library(reactable)
```

The `moviesData` dataset can be displayed with an interative data table.

```{r}
df <- data.frame(
u1 = c(5, 1, 4, NA, NA, NA, 2, 1),
u2 = c(5, NA, 3, 2, 5, NA, 2, 1),
u3 = c(1, 4, 2, 5, 2, NA, 5, 4),
u4 = c(4, 2, NA, 1, 5, 2, 3, 3)
)
reactable(df, defaultPageSize = 4, filterable = TRUE, searchable = TRUE)
```

To then calculate the second user's preference for movie 6, we can calculate the correlations between the second user and the other users, and then use those correlations to predict the second user's rating for movie 6.

```{r}
# Define targets
target_user <- "u2"
target_item <- 6

# 1. Calculate correlations between the target user and all others
correlations <- cor(df, use = "pairwise.complete.obs")[, target_user]
# Remove the target user from the vector so we only have the neighbors
correlations <- correlations[names(correlations) != target_user] 

# 2. Calculate the mean rating for every user
user_means <- colMeans(df, na.rm = TRUE)

# 3. Isolate the neighbors' ratings specifically for Movie 6
# unlist() converts the 1-row data frame into a named numeric vector
neighbor_ratings <- unlist(df[target_item, names(correlations)])

# 4. Filter for neighbors who ACTUALLY rated Movie 6 AND have a valid correlation
valid <- !is.na(neighbor_ratings) & !is.na(correlations) & (correlations > 0)

sims_valid <- correlations[valid]
ratings_valid <- neighbor_ratings[valid]
means_valid <- user_means[names(sims_valid)]

# 5. Apply the collaborative filtering formula
target_mean <- user_means[target_user]

predicted_rating <- target_mean + 
  sum(sims_valid * (ratings_valid - means_valid)) / sum(abs(sims_valid))

# View the result
predicted_rating
```

Thus the predicted rating for user 2 on movie 6 is approximately 2.142857.

## Problem 2:
To use the Pearson correlation to determine wich movie is most similar to movie 1, we can calculate the correlation between movie 1 and all other movies in the dataset.

```{r}
# 1. Transpose df so movies (rows) become columns
movie_matrix <- t(df)

# Assign clear names to the movie columns (Movie 1 to Movie 8)
colnames(movie_matrix) <- paste0("Movie_", 1:8)

# 2. Calculate pairwise correlations between all movies
movie_correlations <- cor(movie_matrix, use = "pairwise.complete.obs")

# 3. Extract correlations for Movie 1 against all other movies (excluding Movie 1)
m1_correlations <- movie_correlations[1, -1]

# 4. Find the movie with the highest correlation
most_similar_movie <- names(which.max(m1_correlations))

# Print results
print(m1_correlations)
cat("The movie most similar to Movie 1 is:", most_similar_movie, "\n")
```

## Problem 3A:

To apply mean imputation grouped according to region to fill in missing values in the column for "Agriculture" and the
column for "Industry" for the `countries` dataset, the following code can be used:

```{r}
# Load the countries dataset
df <- read.csv("http://manctsui.github.io/datasets/countries.csv")


df_grouped <- df  |>
  group_by(Region) |>
  mutate(Agriculture_imputed = ifelse(is.na(Agriculture), mean(Agriculture, na.rm = TRUE), Agriculture),
         Industry_imputed = ifelse(is.na(Industry), mean(Industry, na.rm = TRUE), Industry)) |> ungroup()


df_grouped |> select(Country, Region, Agriculture, Agriculture_imputed, Industry, Industry_imputed)  |> reactable(defaultPageSize = 6, filterable = FALSE, searchable = FALSE)
```

## Problem 3B:
To then find out how many countries were filled in with mean imputation, we can compare the number of missing values before and after the imputation.

```{r}
df_grouped |>
  group_by(Region) |>
  summarise(
    across(
      c(Agriculture, Industry),
      ~ sum(is.na(.))
    ),
    .groups = "drop"
  ) |>
  reactable(defaultPageSize = 11, filterable = FALSE, searchable = FALSE)

```

The NAs before value is equal to the countries filled value, since we are only filling in missing values with mean imputation.

## Problem 4:
We then consider the linear regression model between the variables below
```{r}
df <- data.frame(
temp = c(79, 74, 73, 76, 82, 72, 71, 79, 83, 70, 73, 75, 77),
chirps = c(44, 40, 38, 43, 48, 37, 34, 45, 51, 32, 35, 42, 43)
)
lm_model <- lm(temp ~ chirps, data = df)
summary(lm_model)

```

This summary indicates that the linear regression model is the line Y = 0.7139X + 46.4735, where Y is the temperature and X is the number of chirps. To then manually calculate and compute the linear regression model we can use the following code to define the variables and calculate the slope and intercept:

```{r}
# Define the variables
x <- df$chirps
y <- df$temp

# Calculate the slope (b1) and intercept (b0)
b1 <- cor(x, y) *sd(y) / sd(x)
b0 <- mean(y) - b1 * mean(x)
cat("Slope (b1):", b1, "\n")
cat("Intercept (b0):", b0, "\n")
```

The slope is theoretically derived from the fact that the slope is the covariance of the two variables divided by the variance of the independent variable. The intercept is derived from the fact that the line must pass through the mean of the data points.